---
title: "Sales Revenue Analysis"
author: "David Gabauer"
format: 
  html: default
  pdf: default
editor: visual
---

## Load Required Libraries

```{r}
#| label: load-packages
#| message: false
#| warning: false
#| echo: true

library("gt")
library("readxl")
library("writexl")
library("lubridate")
library("tidyverse")
```

# Data

## Load Dataset {.smaller}

Load `sales.xlsx` and check for potential data quality issues and correct variable types. Each row is uniquely identified by the combination of `customer_id`, `invoice_no`, and `stock_code`. More intuitively, each row represents one item purchased by a customer on a specific invoice.

```{r}
data = read_xlsx("../data/sales.xlsx")
str(data)
```

## Data Cleaning {.smaller}

The data appears free of major quality issues. However, the `quantity` variable is of type character and should be converted to numeric.

```{r}
sales = data |>
  mutate(quantity=parse_number(quantity))
str(sales)
```

## Compute Revenue at the Item-Invoice Level {.smaller}

We now compute the revenue for each item purchased, based on invoice and customer combinations. @tbl-dataset-overview shows the first five observations of the sales dataset.

```{r}
#| label: tbl-dataset-overview
#| tbl-cap: First five rows of the sales dataset
sales = sales |>
  tibble() |>
  mutate(revenue=unit_price*quantity) 

sales |>
  slice(1:5) |>
  gt()
```

# Analysis

## Sales Revenue by Country {.smaller}

@tbl-top-countries shows the top 5 countries by sales revenue. The United Kingdom alone accounts for over 75% of total revenue.

```{r}
#| label: tbl-top-countries
#| tbl-cap: Top 5 countries per sales revenue
sales_country = sales |>
  group_by(country) |>
  summarize(sales = sum(revenue, na.rm = TRUE)) |>
  mutate(percentage = sales / sum(sales) * 100) |>
  arrange(desc(percentage)) |>
  slice(1:5)

sales_country |> 
  gt()
```

@fig-sales_country shows the percentage share of the top 5 countries by total sales revenue.

```{r}
#| label: fig-sales_country
#| fig-height: 7
#| fig-asp: 0.45
#| fig-cap: Bar chart showing the percentage of overall revenue by country
ggplot(sales_country, aes(x=reorder(country, percentage, decreasing=TRUE), y=percentage)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(
    title = "Top 5 Countries by Sales",
    x = "Country",
    y = "Total Sales"
    ) 
```

## Function {.smaller}

The function below calculates daily revenue for a selected country.

```{r}
#| label: create a function
country_daily_revenues = function(data, selected_country='United Kingdom') {
  data$days = as.character(data$days)
  data |>
    filter(country==selected_country) |>
    select(country, year, month, days, revenue) |>
    group_by(days) |>
    summarize(daily_revenue=sum(revenue)) |>
    mutate(days=parse_number(days)) |>
    arrange(days)
}
```

## Daily Revenue in the United Kingdom {.smaller}

@tbl-uk-daily-revenues illustrates daily revenue generated in the United Kingdom during January.

```{r}
#| label: tbl-uk-daily-revenues
#| tbl-cap: Daily revenues in the United Kingdom
country_daily_revenues(sales, selected_country="United Kingdom") |> 
  slice(1:5) |>
  gt()
```

```{r}
sales_uk = country_daily_revenues(sales, selected_country='United Kingdom')

ggplot(sales_uk, aes(x = days, y=daily_revenue)) +
  geom_line(color = "steelblue") +
  labs(
    title = "Daily Revenues Over Days",
    x = "Day of the Year",
    y = "Daily Revenue"
  )
```

## Very Important Customer {.smaller}

@tbl-best-customer lists the customers who generated the highest total revenue.

```{r}
#| label: tbl-best-customer
#| tbl-cap: Revenues by customers
sales |>
  select(country, customer_id, revenue) |>
  filter(!is.na(customer_id)) |>
  group_by(customer_id) |>
  summarize(revenue_per_customer=sum(revenue, na.rm=TRUE)) |>
  arrange(desc(revenue_per_customer)) |>
  slice(1:5) |> 
  gt()
```

## Top Revenue-Generating Invoice {.smaller}

A team member suggested that the customer associated with the invoice generating the highest revenue should receive a promotional gift. This prompted the following analysis.

@tbl-best-invoice shows that the invoice with the highest revenue exceeds the total monthly revenue of the top customer. Your supervisor argued this is not possible and asked for further investigation.

```{r}
#| label: tbl-best-invoice
#| tbl-cap: Revenue by invoice number
sales |>
  select(country, invoice_no, revenue) |>
  filter(!is.na(invoice_no)) |>
  group_by(invoice_no) |>
  summarize(revenue_per_customer=sum(revenue, na.rm=TRUE)) |>
  arrange(desc(revenue_per_customer)) |>
  slice(1:5) |> 
  gt()
```

## Customer of Top Revenue Generating Invoice {.smaller}

Therefore, you need to identify which customer is linked to the highest invoice.

```{r}
sales |>
  filter(invoice_no=="541431") |>
  select(country, customer_id, invoice_no, revenue) |>
  slice(1:5) |>
  gt()
```

Next, you examine the invoice history of the customer. It turns out the customer returned the item. Consequently, you recommend selecting the customer with the highest monthly revenue for the promotion.

```{r}
sales |>
  filter(customer_id=="12346") |>
  select(country, customer_id, invoice_no, revenue) |>
  gt()
```
